---
title: "Bayesian multi-level model for data with structured dispersion"
output:
  html_notebook: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


This post provides code to model variance components in a multi-level setting. This is useful when variances are heteroskedastic and, we suspect, heteroskedasticity is driven by some observed variable(s).

#### Steps: 

1. Load the data. For this demonstration I will simulate data in which 600 individual ($i$) are nested into 200 groups ($j$). $Y$ is the outcome variable and I allow for random intercepts at the group level plus structured dispersion at both group-level and individual-level. In this particular case the variance components are functions of the same variable $z$, but this predictor has opposite effects on each of them.

More formally, the DGP can be described as: $Y = \mathbf{1}_{k} \otimes \mathbf{1}_{n}\mu + \mathbf{I}_{k} \otimes \mathbf{1}_{n} \mathbf{a} + \mathbf{I}_{k} \otimes \mathbf{I}_{n} \mathbf{e}$, where $\mathbf{a}$ and $\mathbf{e}$ are random effects, such that $\mathbf{a} \sim N(\mathbf{0}, \mathbf{\sigma_{a}})$ and $\mathbf{e} \sim N(\mathbf{0}, \mathbf{\sigma_{e}})$. Moreover, both variances are functions of $z$: $\mathbf{\sigma_{a}}=f(\mathbf{z})$ and $\mathbf{\sigma_{e}}=g(\mathbf{z})$.


```{r}

n=3   # Number of individuals by group
k=200 # Number of  groups

# Design matrices for multi-levels

One.k=matrix(rep(1,k))
One.n=matrix(rep(1,n))
I.k=diag(k)
I.n=diag(n)
mu=0

# variance predictor 

z = runif(k,1,5)               # indep variable in variance function at group level
Z = (I.k%x%One.n)%*%z          # indep variable in variance function at individual level 


# An arbitrary function for variance at the group level

THETA_a0 =  -5
THETA_a1  = 1.9
var.a = exp(THETA_a0 + THETA_a1*z + rnorm(k))
sd.a  = sqrt(var.a)


# An arbitrary function for variance at individual level (residual variance)

THETA_e0 =  5
THETA_e1  = -2.7
var.e = exp(THETA_e0 + THETA_e1*Z + rnorm(k*n))
sd.e  = sqrt(var.e)

# Random intercepts
a=rnorm(k,0,sd.a)

# Random error 
e=rnorm(k*n,0,sd.e) 


# Creates Y = mu + random intercept + random error

Y = (One.k%x%One.n)*mu + (I.k%x%One.n)%*%a + (I.k%x%I.n)%*%e   

# Generate indicators. To be used later. 
group=matrix(rep(1:k))  # vector with group indicator for each group
Group=(I.k%x%One.n)%*%group # vector with Group indicator for each individual
```


A simple scatterplot of $Y$ and $z$ should reveal variance structure. 


```{r,echo=FALSE,message=FALSE}
library("tidyverse")

data <- data.frame(y=Y,z=Z)

plot <- data %>% ggplot(aes(x=Z,y=Y)) + geom_point(alpha=0.33, colour="blue") + 
		ylim(-5.5,5.5) + 
		labs(x="Z",y="Y")
print(plot)
```


2. Load ```rstan``` to perform Bayesian estimation via Halmiltonian Monte Carlo. Put the data into a list.


```{r,message=FALSE}

library("rstan")

Data_l1 <- data.frame(Y,Group,Z) # data individual-level
Data_l2 <- data.frame(group,z)   # data group-level 
data    <- list(N=n*k,J=k,y=Data_l1$Y,G=Data_l1$Group,Z=Data_l1$Z, g=Data_l2$group,z=Data_l2$z)

```


3. Write the model in ```stan```.

```{r, message=FALSE}

structured_dispersion <- '

data {
  int<lower=0> J;                   // number of groups 
  int<lower=0> N;                   // number of individuals
  int<lower=1, upper=J>  G[N];      // map individuals to groups
  vector[N] y;                      // outcome variable
  vector[J] z;                      // group level dispersion predictor
  vector[N] Z;                      // individual level dispersion predictor
}


parameters {

  vector[J] a;                           // Random Intercept           
  
  real theta_a0;                         // Intercept model for variance of random intercept
  real theta_e0;                         // Intercept model for variance of residual error
  
  real theta_a1;                         // Slope model for variance of random intercept
  real theta_e1;                         // Slope model for variance of residual error

  vector<lower=0>[J] sigma2_a;           // heterogeneous var for random intercept 

  vector<lower=0>[N] sigma2_e;           // heterogeneous var residual error

}


transformed parameters {
  
  vector[N] y_mu;                        // Random effects

  vector[J] logsigma2_a_mu;              // mean of log-variance for random intercept 
  vector[J] sigma2_a_mu;                 // mean of variance for random intercept 
  vector[J] sigma_a;                     // var for random intercept 

  vector[N] logsigma2_e_mu;              // mean of log-variance residual error
  vector[N] sigma2_e_mu;                 // mean of var residual error
  vector[N] sigma_e;                     // var residual error


  for (i in 1:J) {                       
                                         // Assigns a different sd to each random intercept
    logsigma2_a_mu[i] = theta_a0 + theta_a1*z[i];
    sigma2_a_mu[i]    = exp(logsigma2_a_mu[i]);
    sigma_a[i]        = sqrt(sigma2_a_mu[i]);
  }

  for (i in 1:N) {                    
                                          // Assigns a different sd to each residual error   
    logsigma2_e_mu[i] = theta_e0 + theta_e1*Z[i];
    sigma2_e_mu[i]    = exp(logsigma2_e_mu[i]);
    sigma_e[i]        = sqrt(sigma2_e_mu[i]);

                                          // Assigns a different predicted mean to ech individual depending of group of belonging
    y_mu[i] = a[G[i]];                  
  }


}

model {

  // Priors 

  theta_a0    ~ cauchy(0,2.5);
  theta_e0    ~ cauchy(0,2.5);
  theta_a1    ~ cauchy(0,2.5);
  theta_e1    ~ cauchy(0,2.5);

  // Variances
  sigma2_a ~ lognormal(sigma2_a_mu,0.25);
  sigma2_e ~ lognormal(sigma2_e_mu,0.25);

  // Random intercept 
  a ~ normal(0, sigma_a);

  // Outcome variable
  y ~ normal(y_mu, sigma_e);

}
'

```


4. Fit the model and compare the estimates to the parameters used in the DGP. 

```{r, message=FALSE, results="hide"}
mymodel <- stan(model_code=structured_dispersion, model_name="structured_dispersion", 
iter=2000, pars=c("theta_e0", "theta_e1","theta_a0", "theta_a1"), data=data, chains=3)

```

```{r}
print(mymodel)
```


5. Uset he function```launch_shinystan``` from package ```rstanarm``` to get visual and numerical summaries of model parameters and convergence diagnostics.






