---
title: "Goodness of fit statistics for log-linear models"
output:
  html_notebook: default
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


This code provides (hopefully) useful code to extract goodness of fit statistics for log-linear models fitted via Poisson regression in R. 

1. Start with an arbitrary 5x5 contingency table and put it into dataframe format.

```{r}
	table <- matrix(round(runif(25,0,100),0),5) # creates arbitrary table
	colnames(table) <- paste0("Y_",seq(1:5)) # Column names
	rownames(table) <- paste0("X_",seq(1:5)) # Row names
	df <- data.frame(as.table(table)) # transform table into a dataframe

	head(df)
```

<br>

2. Now, we fit a saturated model and a models of independence on these data.

```{r}
	model.indep <- glm(Freq ~ factor(Var1) + factor(Var2), family=poisson, data=df) # independence model
	model.sat   <- glm(Freq ~ factor(Var1)*factor(Var2), family=poisson, data=df)   # saturated model
```

<br>

3. The next step in standard log-linear analyses is to compare these models using various goodness-of-fit statistics (e.g. G2,AIC,BIC). The following function computes some common statistics from the ```glm``` objects containing the log-linear models.   

```{r}
	gof <- function(freq,model) {
		n    <- sum(freq) 
		df   <- model$df.residual
		q    <- model$df.null - model$df.residual
		G2   <- model$deviance
		D 	 <- (sum(abs(freq-exp(predict(model))))/ (2*n))
		AIC  <- G2 - 2*df
		BIC  <- G2 + log(n)*q
		stats <- c(df,G2,D,AIC,BIC)
		names(stats) <- c("df","G2","D","AIC","BIC")
		return(round(stats,3))
	}
```

The function returns each model's degrees of freedom ($df$), the log-likelihood ratio ($G^2$), dissimilarity index ($D$), Akaike information criterion ($AIC$) and Bayesian information criterion ($BIC$). For example:

```{r}
# Gofs for independence model
gof(df$Freq,model.indep)

# Gofs for saturated model
gof(df$Freq,model.sat)

```

Now it's easy to put them together into a table:

```{r}
	models <- c("model.indep","model.sat")
	table.gof   <- NULL # Empty table


	for (i in models) {
	  tgof <- gof(df$Freq, eval(parse(text = i))) # Computes Goodness of fit statistics for each model
		table.gof <- rbind(table.gof,tgof)  # Piles stats for each model
	}

	rownames(table.gof) <- c("Independence","Saturated")
	print(table.gof) 
```




