---
title: "Lasso regularization for selelection/specification of log-linear models"
output:
  html_notebook: default
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


This post provides code to implement inductive selection/specification of log-linear models for contingency tables via the Lasso. I will exemplify the step-by-step implementation using data on intergenerational occupational mobility for 16 countries.

#### Steps:

1.  Install and load package ```glmnet```, which aperforms Lasso or elastic-net regularization path for generalized linear models. Other packages for data manipulation will be also used.

```{r, message=FALSE}
    library("glmnet")
    library("tidyverse")
    library("modelr")
    library("reshape2")
```

<br>

2. Input the contingency table an turn it into a data frame. I will use the dataset ```hg16``` from the package ```logmult```. This is a cross-classification of subject's occupational status (destination) and his father's occupational status (origin) across 16 countries. 

```{r, message=FALSE}
    # Inpute data and create contingency table as data.frame()
    library("logmult")
    data(hg16)
    table <- ftable(hg16)
    mydata <- as.data.frame(table)
    names(mydata) <- c("origin","destination","country","Freq")
```

This is what the data looks like:

```{r, echo=FALSE}
    print(mydata %>% as_tibble())
```

<br>

3. Next, create the design matrix for the saturated sodel. In this example the saturated models contains 144 parameters. 

```{r}
    # Set reference categories

    mydata$origin      <- relevel(mydata$origin,ref = "Farm")
    mydata$destination <- relevel(mydata$destination,ref = "Farm")
    mydata$country     <- relevel(mydata$country,ref = "Spain")

    # Outcome variable is the frequencies in the contingency table

    y <- mydata$Freq + 0.5 # add small constant to avoid problems with empty cells. 

    # Creates design matrix with all three-way interactions

    formula <- as.formula( ~ .*.*.)
    x <- model.matrix(formula, mydata[,c("origin","destination","country")])
```

<br>

4. Create a vector of penalty factors.

```{r}
    # Penalty factors for weighted penalties. 
    # Here all variables have a penalty of 1, which is equivalent to unweighted penalties. 

    w <- rep(1,dim(x)[2])

    # Unsilence the lines below if you wish to implement the "adaptive lasso" based on ridge estimates

    #ridge.cv <- cv.glmnet(x,y,alpha=0,family='poisson', nfolds = 10)
    #best_ridge_coef <- as.numeric(coef(ridge.cv, s = ridge.cv$lambda.min))[-1] 
    #w <- 1/(abs(best_ridge_coef))^(1)

```

<br>

5. Fit the saturated model using the ```glmnet``` package. In this case you should set the parameter $\alpha=1$ so that you use a Lasso penalty (alternatively, $\alpha=0$ corresponds to a ridge penalty and $\alpha=0.5$ is elastic nets). Because the dependent variable is a vector of frequencies in a contingency table I use a Poisson family with a log-link function. 
    
```{r}
    # "x" is the design matrix for the saturated model 
    # "y" is the  dependent variable
    # "alpha=1" indicates the use of a Lasso penalty
    # "family=poisson" uses a log-link function by default
    # "penalty.factor = w specifies the weight for each variable's penalty
    
    # Fit model
    
    lasso <- glmnet(x,y,alpha=1,family='poisson', penalty.factor = w)

    # Plot the path of coefficients for values of lambda

    plot(lasso,xvar="lambda")
```

<br>

6. Select a value of $\lambda$ that minimizes a measure of cross-validation error. In the case of a Poisson model such measure is based on the Poisson Deviance. For this we use 10-fold cross-validation. Following the advice of the package developers I select the value of $\lambda$ that yields the most regularized model such that the error is within one standard error of the minimum.


```{r}

    # Cross-validation of Lambda

    lasso.cv <- cv.glmnet(x,y,alpha=1,family='poisson', nfolds = 10, penalty.factor = w)

    # Plot cross-validation error for each value of lambda

    plot(lasso.cv)
```

<br>

7. Extract coefficients corresponding to the chosen value of lambda. The vector of coefficients from the Poisson model is most likely sparse. 
 
```{r}
# Extract optimum value of lambda 

opt.lam <- c(lasso.cv$lambda.1se)
lasso.coefs <- coef(lasso.cv, s = opt.lam)
print(lasso.coefs)
```

<br>

8. You can plot the results to easly visualize the margin-free association between origin and destination. 


```{r, echo=FALSE}

# Create predictions for margins and LORative mating. Used in plot 

dummy.model <- lm(Freq ~ origin + destination + country, data=mydata)
new_x <- mydata %>% data_grid(origin,destination,country,.model=dummy.model) %>% model.matrix(formula, .)


# full prediction
predictions <- cbind(mydata%>% data_grid(origin,destination,country,.model=dummy.model), predict(lasso.cv, new_x, s=opt.lam)) %>%
  as_tibble() %>% rename(pred = `1`) 


# Intercept
intercept <- predictions %>% filter(origin=="Farm", destination=="Farm", country=="Spain") %>% summarise(pred) %>% as.numeric()


# margins
predictions <- predictions %>% mutate(pred = pred - intercept) # remove intercept

predictions_country     <- predictions %>% filter(origin=="Farm", destination=="Farm")  %>% rename(margin_country=pred) %>% select(country,margin_country)
predictions_origin      <- predictions %>% filter(country=="Spain", destination=="Farm") %>% rename(margin_origin=pred) %>% select(origin,margin_origin)
predictions_destination <- predictions %>% filter(country=="Spain", origin=="Farm") %>% rename(margin_destination=pred) %>% select(destination,margin_destination)


# match
predictions <- predictions %>% left_join(predictions_country, by="country")
predictions <- predictions %>% left_join(predictions_origin, by="origin")
predictions <- predictions %>% left_join(predictions_destination, by="destination")


# margins by country
predictions_country_origin <- predictions %>% filter(origin!="Farm",country!="Spain",destination=="Farm") %>%
  rename(margin_country_origin=pred) %>% mutate(margin_country_origin = margin_country_origin - (margin_country + margin_origin )) %>%
  select(country,origin,margin_country_origin)

predictions_country_destination <- predictions %>% filter(origin=="Farm",country!="Spain",destination!="Farm") %>%
  rename(margin_country_destination=pred) %>% mutate(margin_country_destination = margin_country_destination - (margin_country + margin_destination )) %>% 
  select(country,destination,margin_country_destination)

predictions <- predictions %>% left_join(predictions_country_origin, by=c("country","origin")) %>%  replace_na(list(margin_country_origin = 0))
predictions <- predictions %>% left_join(predictions_country_destination, by=c("country","destination")) %>%  replace_na(list(margin_country_destination = 0))


# computes margin-free log-odd rations (LORs)
predictions <- predictions %>% 
  mutate(LOR = pred - (margin_country + margin_origin + margin_destination + margin_country_origin + margin_country_destination) ) 
levels.origin      <- c("Farm","Blue Collar","White Collar")
levels.destination <- c("Farm","Blue Collar","White Collar")
levels.country     <- levels(mydata$country)


plot <- predictions %>% 
  ggplot(aes(y=factor(origin, levels = rev(levels.origin)),
             x=factor(destination, levels = levels.destination))) + facet_wrap( ~ country) + geom_raster(aes(fill= LOR)) +
  scale_fill_gradientn(limits=c(-3.5,3.5), colours=c("red","white","blue")) +
  labs(y="Father's occupation", x= "Children's occupation", colour="") +
  theme_bw() + theme(axis.text.x = element_text(size=9, angle=45, vjust=-1, hjust=0),
                     axis.text.y = element_text(size=9, angle=0),
                     plot.title= element_text(size=11)) +
  scale_x_discrete(position="top")


print(plot)


```


