---
title: "Simple Linear Regression Model"
author: "Limeng(Momo) Xie"
date: "4/23/2020"
output:
  html_document:
    keep_md: true
---

Simple liner regression model 
Yi - a + bXi + ei 
X: the independent variable. We assume X is fixed 
Y: the dependenat variable, random 
e: error term. We assume that ei are the independent and have the same probability distribution normal (0, sigma^2)

The probablity distribution of Yi is Normal(a+bXi, sigma^2)
```{r, echo = FALSE}
knitr::opts_chunk$set(
  fig.path = "README_figs/README-"
)
# require ggplot
require(ggplot2)
# read the file 
data <- read.delim("~/Desktop/BINF8441_Course/poverty.txt")
head(data)
summary(data)
# y = year 2002 birth rate per 1000 females 15 to 17 years old and x = poverty rate, which is the percent of the state’s population living in households with incomes below the federally defined poverty level

# (1) scatter plot 
ggplot(data=data,aes(x=PovPct, y=Brth15to17))+
    geom_point()+
    theme_classic()+
    labs(title="Birth rate per 1000 females (15 to 17 years old) by poverty rate", x="Poverty Rate", y="Birth rate per 1000 females")
```

**we can see there is a positive relationship between poverty rate and Birth rate per 1000 females (15 to 17 years old)** 

b) we can test there is a liner relationship of Y and X  using lm function 
```{r}
x = data$PovPct
y = data$Brth15to17
result = lm(y~x)
summary(result)
```
**the result shows that the coeffient b(slope) is 1.3733 >0 with a significant level less than 0.05, which indicates that Y and X is positively correlated**