---
title: "Lab4 Machine Learning in R"
author: "Yongjun Zhang, Ph.D."
institute: "Department of Sociology and IACS, Stony Brook Unversity"
date: "02/14/2022"
output:
  rmdformats::readthedown:
    highlight: pygments
--- = =
---

```{=html}
<style type="text/css">
p{ /* Normal  */
   font-size: 18px;
}
body{ /* Normal  */
   font-size: 18px;
}
td {  /* Table  */
   font-size: 14px;
}
h1 { /* Header 1 */
 font-size: 32px;
}
h2 { /* Header 2 */
 font-size: 26px;
}
h3 { /* Header 3 */
 font-size: 22px;
}
code.r{ /* Code block */
  font-size: 14px;
}
pre { /* Code block */
  font-size: 14px
}
</style>
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Learning Objects

This tutorial aims to introduce some basic ways to implement ml in R.

1. Learn how to use R caret package to train a supervised ml model.

2. Learn how to use R keras to train a basic neural network mode.

3. Use SSA US baby names to train a model to predict gender


## Use caret package to train classic ML model

Max Kuhn has detailed documentation related to how use caret pacakge to train and tune models. You can check here for more details: <https://topepo.github.io/caret/>.

You can also get his book on Applied Predicting Modeling: <http://appliedpredictivemodeling.com/>

In this tutorial, we are going to rely on Max Kuhn's caret package to train a simple classifier to predict gender. 

> The caret package (short for Classification And REgression Training) is a set of functions that attempt to streamline the process for creating predictive models. The package contains tools for: data splitting; pre-processing; feature selection; model tuning using resampling; variable importance estimation

Of course, there are other packages that you can use to train your own models, but I prefer using caret as it is one of the most popular in our community. 

Here our goal is to use data from SSA to train a gender classifier using name features.

### Data Preparation

> You can download ssa baby names via here <https://catalog.data.gov/dataset/baby-names-from-social-security-card-applications-national-data>

> I have also used the following codes to preprocess the data. I have generated aggregated name counts by gender and then define our target outcome as "female if over 50 of women use that name." I have also created four variables: first letter, first two letters, last letter, and laster two letters. You can directly use the RData file, <https://yongjunzhang.com/files/ssa_baby_names.RData>. 

> Of course, you can take a more rigorous way to define the gender of names.

```{r}
files <- list.files(path = "./names/",pattern = ".txt",full.names = TRUE)
files

# let us read the first file
library(tidyverse)
# let us say we want to define a read_us_baby function
readSsaBabyNames <- function(file,...){
  require(tidyverse)
  data <- read_csv(file,col_names = FALSE,show_col_types = FALSE) %>% 
    mutate(year=str_replace_all(file,"[^0-9]",""))
  colnames(data) <- c("names","sex","count","year")
  return(data)
}

dat_year <- map_dfr(files, readSsaBabyNames)

# we only use names after 1970 and used by at least 10
dat_all <- dat_year %>%
  filter(year>1970) %>% 
  group_by(names,sex) %>% 
  summarise(count=sum(count)) %>% 
  # we only keep names used by at least 10 
  filter(count>10) %>% 
  pivot_wider(names_from = "sex",values_from="count") %>% 
  replace_na(list(F=0,M=0)) %>% 
  mutate(female=(F/(F+M)>.5)*1) %>% 
  mutate(flt1= str_extract(names,"^.")%>% tolower,
         flt2= str_extract(names,"^.{2}") %>% tolower,
         llt1= str_extract(names,".$")%>% tolower,
         llt2= str_extract(names,".{2}$")%>% tolower
         )
  
  
save(dat_year,dat_all,file="./ssa_baby_names.RData")
```

### Split our data into train and test data

Before we further split our data, let us take a look at data first.

```{r}

top_letters <- dat_all %>% 
  group_by(flt1) %>% 
  summarise(fl1=n()) %>% 
  top_n(n=5) %>% 
  bind_cols(
    dat_all %>% 
      group_by(flt2) %>% 
      summarise(fl2=n()) %>% 
      top_n(n=5)
  ) %>% 
  bind_cols(
    dat_all %>% 
      group_by(llt1) %>% 
      summarise(ll1=n()) %>% 
      top_n(n=5)
  ) %>% 
  bind_cols(
    dat_all %>% 
      group_by(llt2) %>% 
      summarise(ll2=n()) %>% 
      top_n(n=5)
  )

knitr::kable(top_letters)

```

We are going to create a series of features based on first/first two and last/last two letters to predict gender of names. But in your own training, you should do some feature engineering and choose those most informative features.

```{r}
library(caret)
#create dummies for last letter, last two letters, fist letter, and first two letters of names
fllt_d=predict(dummyVars(~llt1+llt2+flt1+flt2,data=dat_all),newdata=dat_all)%>%as.data.frame()
# This will create a over 900 vars, and some of them are near zero variance, so we need to get rid of it.
#identify near zero variance variables
nzv_test <- nearZeroVar(fllt_d,freqCut = 95/5, uniqueCut = 10, saveMetrics= TRUE,allowParallel = TRUE)
nzv <- nearZeroVar(fllt_d,freqCut = 95/5, uniqueCut = 10,allowParallel = TRUE)
fllt_nnzv<- fllt_d[, -nzv]
#create the full datasets with dummies
df=dat_all %>% 
  ungroup %>% 
  select(names,female) %>% 
  bind_cols(fllt_nnzv) %>% 
  filter(!is.na(female)) %>% 
  mutate(female=ifelse(female==1,"Y","N") %>% as.factor())
```

Let us finally split our data into train and test

```{r}
inTrain <- createDataPartition(
  y = df$female,
  ## the outcome data are needed
  p = .75,
  ## The percentage of data in the
  ## training set
  list = FALSE
)

train <- df[ inTrain,]
test  <- df[-inTrain,]

nrow(train)
#> [1] 157
nrow(test)
library(gdata)
keep(dat_all,df,train,test,sure=TRUE)

```

### Let us train a xgboost model to predict gender

You can check caret package's boosting methods here <https://topepo.github.io/caret/train-models-by-tag.html#boosting>. Of course, you can fine-tune hyper-parameters, but we use the default ones.

```{r}

# K folds cross validation
# try parallel computing
library(doParallel)
library(xgboost)
cl <- makePSOCKcluster(3)
registerDoParallel(cl)

grid_default <- expand.grid(
  nrounds = 100,
  max_depth = 6,
  eta = 0.3,
  gamma = 0,
  colsample_bytree = 1,
  min_child_weight = 1,
  subsample = 1
)

train_control <- caret::trainControl(
  method = "cv",
  number = 3,
  verboseIter = FALSE, 
  allowParallel = TRUE 
)

xgb_base <- caret::train(
  female~., 
  data=train %>%  dplyr::select(-c(names)),
  trControl = train_control,
  tuneGrid = grid_default,
  method = "xgbTree",
  verbose = TRUE
)

stopCluster(cl)

save(xgb_base,file="./xgb_base.RData")

```

### Let us check model performance, get the comfusion matrix firrst

```{r}
load("./xgb_base.RData")
# check cf matrix
xgb_base
confusionMatrix(xgb_base)

```


### Let us check the most informative features


```{r}
xgbImp <- varImp(xgb_base,scale = TRUE)
plot(xgbImp, top = 5)
```

### Test model performance on test set

```{r}

# predict name_remove
test_df <- test %>% dplyr::select(-c(names,female))
test_pred <- predict(xgb_base,newdata=test_df) %>% as.data.frame()
colnames(test_pred) <- "xgb_female"

data <- test %>% select(names,female) %>% bind_cols(test_pred)

data %>% 
  write_csv("./test_pred.csv",na="")

# check confusion matrix

confusionMatrix(test_pred$xgb_female,test$female)
```

## Code Challenge 1

> Replicate the ML predicting baby names' gender using top 5 most frequent first/last and first-two/last two letters instead of all dummies

> Try other models, including support vector machine, penalized logit model, and naive bayes; You can also try random forest, but it takes a while to finish.

> Generate your final prediction based on svm, logit, and nb methods (using the majority vote)

> Save a csv file and report the confusion matrix

> You codes should be written in Rmarkdown file and genearte a pdf file.

> Email me before next Monday class.

