%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Medium Length Professional CV
% LaTeX Template
% Version 2.0 (8/5/13)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% Original author:
% Trey Hunner (http://www.treyhunner.com/)
%
% Important note:
% This template requires the resume.cls file to be in the same directory as the
% .tex file. The resume.cls file provides the resume style used for structuring the
% document.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass{resume} % Use the custom resume.cls style

\usepackage{hyperref}
\usepackage[left=0.75in,top=0.6in,right=0.75in,bottom=0.6in]{geometry} % Document margins
\newcommand{\tab}[1]{\hspace{.2667\textwidth}\rlap{#1}}
\newcommand{\itab}[1]{\hspace{0em}\rlap{#1}}
\name{Nicholas Sale} % Your name
\address{Dial House, The Street, Gosfield, Essex CO9 1TU} % Your address
\address{New College, Holywell Street, Oxford OX1 3BN} % Your secondary addess (optional)
\address{nicholas.j.sale@gmail.com \\ 07842 118883} % Your phone number and email

\begin{document}

%----------------------------------------------------------------------------------------
%	EDUCATION SECTION
%----------------------------------------------------------------------------------------

\begin{rSection}{Education}

{\bf University of Oxford} \hfill {\em October 2015 - Present} 
\\ MMathCompSci Mathematics and Computer Science
\\ \\
\itab{\textbf{Relevant Courses Taken}} \tab{}  \tab{\textbf{Being Taken This Year}}
\\ \itab{Topology} \tab{} \tab{Algebraic Topology} 
\\ \itab{Topology \& Groups} \tab{} \tab{Computational Algebraic Topology} 
\\ \itab{Rings \& Modules} \tab{}  \tab{Homological Algebra} 
\\ \itab{Machine Learning} \tab{} \tab{Categories, Proofs and Processes}
\\ \itab{Introduction to Representation Theory} \tab{} \tab{Computational Learning Theory}
\\ \itab{Knowledge Representation \& Reasoning} \tab{} \tab{Mathematics and Data Science for Development}

\end{rSection}

%----------------------------------------------------------------------------------------
%	RESEARCH EXPERIENCE SECTION
%----------------------------------------------------------------------------------------

\begin{rSection}{Research Experience}
	
\begin{rSubsection}{Applied Research Summer Placement}{\em July 2018 - September 2018}{UK Civil Service}{Civil service references available on request}
	\item 11-week placement within a small team working on statistical analyses of the impact of certain government cyber-security projects, where I was tasked with the open-ended goal of researching how machine learning and other data science techniques could be applied to aid the team with data annotation.
	\item Performed and wrote up experiments to support judgements on the efficacy and utility of different methods for automating this task, applying techniques from natural language processing, statistical learning, and topological data analysis.
	\item Presented my work and an introduction to topological data analysis in a seminar for the data science community within the government department.
	\item Also participated in the initial literature review for a new project on embedding internet-scale network graphs, in collaboration with the Alan Turing Institute.
\end{rSubsection}

%------------------------------------------------

\begin{rSubsection}{Applied Research Summer Placement}{\em July 2017 - September 2017}{UK Civil Service}{}
	\item 11-week placement within a government applied research department working on developing internet-related capability, where I was tasked with determining the feasibility of using data science techniques to identify certain types of network devices based on limited information about their traffic.
	\item Evaluated different models for network devices and produced visualisations to aid feature engineering, before experimenting with different statistical learning techniques for the classification.
	\item Presented my work at the kick-off meeting for a workshop on analysing network traffic.
\end{rSubsection}

%------------------------------------------------

\begin{rSubsection}{Summer School Individual Research Project}{\em July 2016 - September 2016}{UK Civil Service}{}
	\item Undertook a short project investigating the use of sheaves and sheaf cohomology in data science, following Michael Robinson's 'Tutorial on Sheaves in Data Analytics', and writing a short Haskell script to perform calculations.
	\item See \url{https://github.com/NickSale} for code.
\end{rSubsection}


%------------------------------------------------

\begin{rSubsection}{Work Experience}{}{BAE Systems, Great Baddow \hfill July 2013 \\ Microsoft Research, Cambridge}{February 2014}
	\item Spent two weeks at BAE Systems and a week at Microsoft Research learning about industrial research.
\end{rSubsection}

\end{rSection}

%----------------------------------------------------------------------------------------
%	TECHNICAL SKILLS SECTION
%----------------------------------------------------------------------------------------

\begin{rSection}{Technical Skills}

\begin{tabular}{ @{} >{\bfseries}l @{\hspace{6ex}} l }
Programming Languages &  Experienced: Python, Java, C$^{\sharp}$ \quad Familiar: C++, Scala, PHP, Haskell\\
Software \& Tools & \LaTeX, scikit-learn, pandas, Spark (parallel cluster computing) \\
\end{tabular}

\item Developed my technical skills and knowledge during an intensive 10-week summer school in 2016 which covered advanced topics in computer security, including some relevant applications of data science.

\item Worked on a 6-month engineering project as part of a team for the \textit{Engineering Education Scheme} in 2014, working with Selex ES to design and build an autonomous vehicle. I took lead on researching, designing, and implementing the sensing, geometry representation and path-planning aspects of the project.

\end{rSection}

%	ACADEMIC ACHIEVEMENTS SECTION
%----------------------------------------------------------------------------------------

\begin{rSection}{Academic Achievements} \itemsep -2pt

\begin{rSubsection}{New College Undergraduate Scholarship}{\em October 2016 - Present}{}{New College, Oxford}
	\item Awarded for examination performance.
\end{rSubsection}

%------------------------------------------------

\begin{rSubsection}{CyberFirst Bursary Scheme}{\em October 2015 - Present}{}{}
	\item A government and industry-backed initiative aiming to improve the UK's cybersecurity capability by supporting selected undergraduate students.
\end{rSubsection}

%------------------------------------------------

\begin{rSubsection}{GResearch Computer Science Project Prize}{\em 2017}{}{New College, Oxford}
	\item Awarded for innovation in producing an automatic music composition program for a group project.
\end{rSubsection}

%------------------------------------------------

\begin{rSubsection}{Arkwright Scholarship}{\em September 2013 - September 2015}{}{}
	\item A prestigious engineering scholarship awarded after examination and interview.
\end{rSubsection}

\end{rSection}

%	POSITIONS OF RESPONSIBILITY SECTION
%----------------------------------------------------------------------------------------

\begin{rSection}{Positions of Responsibility}

\begin{rSubsectionNoContent}{New College Boat Club Committee}{\em New College, Oxford}{President 2018/19, Secretary 2017/18, Lower Boats Captain 2016/17, W3 Coach 2018}{}
\end{rSubsectionNoContent}

%------------------------------------------------

\begin{rSubsectionNoContent}{Bar Sports Captaincies}{\em New College, Oxford}{Darts Captain 2017/18, Pool Captain 2016/17}{}
\end{rSubsectionNoContent}

%------------------------------------------------

\begin{rSubsectionNoContent}{Oxford University Roleplaying Society Committee}{\em October 2016 - October 2018}{Netrep (IT Secretary)}{}
\end{rSubsectionNoContent}

\end{rSection}

%	EXTRA-CURRICULAR SECTION
%----------------------------------------------------------------------------------------

\begin{rSection}{Extra-Curricular} \itemsep -3pt
\item I row, play bass guitar, and recently completed my Duke of Edinburgh Gold award.

\end{rSection}

\end{document}